Risk Perceptions for Wearable Devices

### 2015-12-11 paper.tex
removed commentedout sections (I have git anyway)
finished rough draft

\begin{abstract}
\indent\indent Wearable devices, or ``wearables,'' bring great benefits but also potential risks that could expose users' activities without their awareness or consent. In this paper, we report findings from the first large-scale survey conducted to investigate user security and privacy concerns regarding wearables. We surveyed 1,782 Internet users in order to identify risks that are particularly concerning to them; these risks are inspired by the sensor inputs and applications of popular wearable technologies. During this experiment, our questions controlled for the effects of what data was being accessed and with whom it was being shared. We also investigated how these emergent threats compared to existent mobile threats, how upcoming capabilities and artifacts compared to existing technologies, and how users ranked technical and non-technical concerns to sketch a concrete and broad view of the wearable device landscape. We hope that this work will inform the design of future user notification, permission management, and access control schemes for wearables.
\end{abstract}

@@

Wearable devices enable many benefits, ranging from interaction with virtual objects in an augmented reality world to healthier, fitness-data inspired lifestyles. However, wearable devices also bring new potential privacy and security risks that could expose users' activities without their awareness or consent. Although wearable devices are still in their infancy, we have already seen manifestations of these risks. Fitbit's default privacy settings inadvertently exposed information about some of their users' sexual activity~\cite{Fitbit}. Public discomfort toward facial recognition caused Google to prohibit Google Glass applications from using facial recognition~\cite{GlassDetection}, but still resulted in tech hate crimes against its users ~\cite{1_russell_2014, 16_gross_2014}. Google Glass has since been discontinued.

For smartphones, security and privacy risks are generally addressed by communicating data capture to users. However, many users are habituated to these notifications, because they see them frequently, often for things that they do not care about~\cite{felt2012android}. Once habituated to seemingly benign privacy and security warnings, users tend to ignore more sensitive warnings that are similarly designed~\cite{Egelman08}.

@@

The goal of this work is to shape the still-malleable future of wearable platforms and interaction models, with research on user-centric concerns. To our knowledge, this is the first large-scale study to investigate user security and privacy concerns for wearable devices. Our survey of 1,782 Internet users contributes the following: %\\[-0.8cm]

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item We report how 72 types of data likely to be captured by wearable devices were perceived by our participants and rank them by relevance. 
\item We repeat this across 4 types of recipients to also illustrate the contribution of the data recipient to the overall perceived risk. 
\item We sketch a landscape of users' self-reported concerns regarding wearable devices, spanning concerns outside of security and privacy. 
%\item We compare emergent risks with existing risks and find that participants perceive risks similarly to physical risks---for instance, facial detection was perceived as risky as using a lawnmower.
\end{itemize}

\section{Methodology}
To obtain a comprehensive list of possible risks that wearable devices might present in the future, we examined the sensors, capabilities, permissions, and applications of the most popular wearable devices on the market. At the time of this study (August 2014) the most popular wearable devices included the Fitbit fitness tracker, which continuously monitors heartbeat, steps taken, and sleep patterns;
%~\cite{6_fitbit_2014, 7_time_2014}
the Pebble smartwatch, which can take pictures, send texts, show notifications from online, and push notifications to services; %~\cite{pebble_smartwatch_2014, 9_verge_2014, 10_readwrite_2014}
and Google Glass, which can take pictures, record video, and perform a subset Internet-based tasks such as search, reading emails, etc. %~\cite{11_wikipedia_2015, 12_turi_2014}. 
These devices' capabilities and requested permissions were our inspiration to develop a list of possible security and privacy risks that users will encounter.

We designed a survey to gauge the relevancy of all these possible risks.
Our survey contained two main sections. A set of questions presented participants with several scenarios---something undesirable that might happen with their wearable device---and asked them to rate their level of concern if each scenario were to happen. This was intended to elicit their perception of the severity and impact of the risk.
The format of this section was based on Felt {\it et al.}'s study of user perceptions of security and privacy risks with mobile devices~\cite{Felt}. Another set asked participants to compare the risks and benefits of wearable technologies to those of better-understood technologies, following the same methodology from Fischhoff {\it et al.}'s seminal study in risk perception~\cite{Fischhoff}.
%Finally, we collected demographic information, which included a privacy concern scale and whether participants owned any wearable devices.

\subsection{Related Work}
{\color{red} some intro here}\\

\noindent {\bfseries Ubiquitous Sensing}

We are rapidly moving towards a world of ubiquitous sensing and data capture, with ensuing privacy challenges~\cite{abowd2000charting,palen2003unpacking,camp2000internet}. Roesner {\it et al.}\ urge the community to address potential concerns for wearable devices before the technologies become widespread~\cite{roesner2014security} and explore the unique and difficult problems these devices present in terms for law and policy~\cite{roesner2014augmented}.

Many researchers have worked to study how privacy can be preserved in the presence of ubiquitous devices. Examples of such efforts include frameworks to design for privacy~\cite{bellotti1993design,camp2003designing,langheinrich2001privacy}, protocols for anonymous communication~\cite{cornelius2008anonysense}, evaluation metrics for privacy~\cite{scholtz2004toward}, and privacy models~\cite{hong2004privacy, jiang2002approximate}. Our work aims to augment works like these with an understanding of what privacy means to the end user. \\

\noindent {\bfseries Lessons form Smartphones}

Smartphones allow applications to access new types of data. While this tends to benefit users, they do not think of the privacy implications. Lindqvist {\it et al.}\  studied a popular location tracking application, and found that people share their location information for gaming, signaling availability to friends, without concern for the privacy implications of broadcasting that information~\cite{lindqvist2011m}. There are still many unresolved concerns such as the opaqueness that prevents users from fully understanding how applications are using their data or rogue applications inappropriately accessing data~\cite{1_kane_2010, zhou2011taming}.

Felt \etal previously studied the security concerns of smartphone users by conducting a large-scale online survey~\cite{Felt}. Their survey asked 3,115 smartphone users about 99 risk scenarios. Participants were asked how upset they would be if a certain action occurred without their permission. Participants rated each situation on a Likert scale ranging from ``indifferent (1)'' to ``very upset (5).'' Our methodology closely follows that study, but with scenarios chosen to shed light on the security and privacy risks of wearable devices.

Tsai {\it et al.}\ found that when mobile users get feedback about releasing data, such as who has viewed location information, it greatly impacts future behaviors~\cite{Tsai2009}. Although this type of feedback is not provided to smartphone users, there is potential for impacting wearables users and shaping their behaviors so that they keep users safe. \\

\noindent {\bfseries User Perceptions}

While risk communication for the physical world has been examined for several decades (e.g.,~\cite{Fischhoff,Morgan2001}), research into effectively communicating computer-based risks has only recently been researched. For example, both Garg {\it et al.}\ and Blythe {\it et al.}\ show that due to varying perceptions and abilities that correlate with demographic factors, computer-based risk communication should employ some degree of demographic targeting~\cite{Garg2012,Blythe2011}. While this work is likely applicable to wearable computing risk communication, we believe that a better understanding of users' risk perceptions in this domain is warranted, prior to examining risk communication.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item 2 reading comprehension questions
\item 6 questions about wearable computing scenarios 
\item 2 questions about smartphone scenarios 
\item 2 Fischhoff-style risk/benefit questions 
\item 15 demographic questions %\\[-.8cm]
\end{itemize}

We randomized the order participants saw sections of the survey (with the exception of the comprehension and demographic questions, which were always first and last, respectively), as well as the order of questions in each section.

\subsubsection{Comprehension Questions}
Because participants might be biased to specific companies (e.g., visceral reactions to Google Glass based on popular media stories), we based our questions on a fictitious wearable. Thus, the beginning of the survey introduced participants to the ``Cubetastic3000,'' which was the basis for all questions on wearables risks. We highlighted the capabilities of this device and described use cases:

\begin{quotation}
{\it Imagine that you are the proud owner of the Cubetastic3000, a new, high-tech computing device designed to be worn on your head. Imagine also that you wear this device all the time, because it is very lightweight, durable, and convenient.

The Cubetastic3000 has the capability to capture video, photos, audio, and biometrics (biological data about you, such as heart rate). Just like other devices today, you can install third-party applications from an app store, and these applications can use the information that the Cubetastic3000 captures.

With a wide range of applications and capabilities, your device can do all sorts of things, such as:\\

@@

Since this device is the basis for many of our questions, we ensured that participants had understood its capabilities by asking them two multiple-choice comprehension questions. We filtered out responses from participants who could not answer these questions.

\subsubsection{Wearable Scenarios}
We presented scenarios involving data captured by the Cubetastic3000 and asked participants to rate how upset they would be if a particular type of data (e.g., video, audio, name, etc.) was shared without permission with a particular data recipient (see Figure \ref{fig:prompt}). Responses were reported on a 5-point Likert scale (from ``indifferent'' to ``very upset''), following Felt \etal\cite{Felt}. Questions were of the form:

@@

We combined 72 data types (<data>) with 4 data recipients (<recipient>) to form an initial pool of 288 questions: %\\[-.8cm]

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{images/prompt.pdf}
	\caption{An example of a wearable scenario question participants saw while taking the survey.}
	\label{fig:prompt}
\end{figure}

\begin{packed_item}
\item Your work contacts
\item Your friends
\item The public
\item The app's server (but didn't share it with anyone else) %\\[-.8cm]
\end{packed_item}

The purpose for using this question format was to determine how upset participants would be if data were inappropriately shared, and the extent to which their reactions were based on the data type and recipient. Each participant answered 6 questions that were randomly drawn from this pool. %293: the 288 described here, plus 5 that we describe in the next section (Section \ref{sec:smartphones}).

%\subsubsection{Smartphone Scenarios}
%\label{sec:smartphones}
%We presented participants with a second set of scenarios to control for the type of device being used. Rather than using the previous pool of 288 <data> and <recipient> combinations, we selected 5 scenarios that Felt \etal found least and most concerning to their participants~\cite{Felt}:
%
%\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
%\item {\it How would you feel if an app on your <device> vibrated your phone without asking you first?}
%\item {\it How would you feel if an app on your <device> connected to a Bluetooth device (like a headset) without asking you first?}
%\item {\it How would you feel if an app on your <device> un-muted a phone call without asking you first?}
%\item {\it How would you feel if an app on your <device> took screenshots when you were using other apps, without asking you first?}
%\item {\it How would you feel if an app on your <device> sent premium (they cost money) calls or text messages, without asking you first?}
%\end{enumerate}
%
%
%In the previously described section of our survey, <device> was set to ``Cubetastic3000'' and not every participant received one of these questions (i.e., these 5 questions were among the pool of 293 questions from which participants were randomly assigned 6). In the separate smartphone section of the survey, every participant received exactly two of these questions, where <device> was set to ``smartphone.'' This allowed us to perform controlled comparisons based on whether the same misbehavior was occurring on a smartphone (i.e., a better understood device) or the Cubetastic3000 (i.e., a fictitious wearable device).

%\subsubsection{Risk and Benefit Assessment}
%In addition to investigating reactions to particular scenarios, we examined broad perceptions of new technologies and how those compared to perceptions of other understood technologies. We modeled this section after a seminal risk perception study by Fischhoff \etal\cite{Fischhoff}, in which participants ranked technologies by their relative risk and benefit to society. We ask participants to perform this exercise for 4 technologies previously examined by Fischhoff {\it et al.}: handguns, motorcycles, lawnmowers, and electricity.  These technologies were chosen to span varying levels of risks and benefits.
%
%Our goal was to ask about familiar technologies such as the Internet, general and specific wearable artifacts, and a range of new capabilities. To do this, we asked participants to evaluate one of 20 technologies relevant to wearables along 4 previously studied technologies. The 20 technologies were: Internet, email, laptops, smartphones, smart watches, fitness trackers, Google Glass, Cubetastic3000, discrete camera, discrete microphone, facial recognition, facial detection, voice recognition, voice-based emotion detection, location tracking, speech-to-text, language detection, heart rate detection, age detection, and gender detection.
%
%To parallel Fischhoff {\it et al.}'s risk perception study, we gave our participants a similar prompt to numerically express the perceived gross risk/gross benefit over a long period of time for all parties involved. We randomized whether they performed the ranking for risks or benefits first. The prompt is listed in Appendix \ref{sec:prompt}. The question format was as follows:
%
%\begin{quotation}
%{\it \noindent Fill in your <risk/benefit> numbers for the following:\\
%
%\noindent Handguns: \_\_\_\_\_\_\_ \\
%Motorcycles: \_\_\_\_\_\_\_\\
%Lawnmowers: \_\_\_\_\_\_\_\\
%<Wearable Technology>: \_\_\_\_\_\_\_\\
%Electricity: \_\_\_\_\_\_\_\\ }
%\end{quotation}

\subsubsection{Additional Questions}
The exit portion of the survey contained demographic questions asking for age, gender, and education. We then asked about wearables ownership so we could control for prior exposure. An open-ended question on what would be the most likely risks associated with wearable devices was included to catch user concerns more broadly. To avoid biasing the open-ended question, we asked before concluding with the 10-question Internet Users' Information Privacy Concerns (IUIPC) index~\cite{malhotra2004internet}, which was included so we could control for participants' general privacy attitudes.

@@

%We followed Fischhoff {\it et al.}'s methodology and did not normalize the numerical responses. Rather, we report medians and quartiles, which are not impacted by outliers. For the open-ended question at the end (i.e., additional privacy concerns), two researchers independently coded 1,782 responses, with an initial agreement rate of 89.7\%. The researchers discussed and resolved any disagreements so that the final codings reflect unanimous agreement.

\section{Results}
In this section, we present participants' responses to the various data-sharing scenarios, and discuss how and which various factors contributed to their risk perceptions. We also discuss participants' risk/benefit assessment of various new technologies relative to well-established technologies. We conclude the section with participants' self-reported concerns about the biggest risks in owning wearable devices.

%\subsection{Concern Factors}
%Many factors can potentially impact participants' concern levels:  the data being shared, with whom the data is shared, and device in question. We analyze each factor individually, as well as present a statistical model of participants' concerns as a function of the above factors and additional demographic traits.

\subsubsection{Data Type}
\label{sec:datatypes}

Based on our statistical models (later reported in Section \ref{sec:regression}), we observe that the largest effect on participants' VURs stemmed from the data being shared; with whom the data is shared and which device the user is interacting with had weaker impacts on overall VURs. The most and least concerning data types are listed in Table \ref{top10-table}, and the full list can be seen in Table \ref{full-vur-table} in Appendix \ref{sec:concerns-appendix}.

@@

Although certain data is considered unanimously upsetting to have shared, it is interesting to note that no data was considered unanimously non-upsetting to have shared, nor were there any data types that evoked strong disagreement between participants (i.e., bimodal). Generally, the average concern magnitude was inversely correlated with the standard deviation, which suggests the presence of ceiling effects for the most concerning data types. For the complete ranked list of data types in this study, see Appendix \ref{sec:concerns-appendix}.

\begin{table*}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | r |c |}
\hline
Rank & Data & VUR & $\sigma$ & Distribution \\
\hline
1 & video of you unclothed & 95.97\% & 0.31 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouunclothedcombined} \\
2 & bank account information & 95.91\% & 0.35 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourbankaccountinformationcombined}  \\
3 & social security number & 94.84\% & 0.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyoursocialsecuritynumbercombined}\\
4 & video entering in a PIN at an ATM & 92.67\% & 0.47 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouenteringinyourPINatanATMcombined}\\
5 & photo of you unclothed & 92.59\% & 0.46 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookaphotoofyouunclothedcombined}\\
6 & photo of you that is very embarrassing & 91.39\% & 0.55 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingphotoofyoudoingsomethingembarrassingcombined}\\
7 & username and password for websites & 89.55\% & 0.62 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourusernameandpasswordforwebsitescombined}\\
8 & credit card information & 88.98\% & 0.56 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourcreditcardinformationcombined}\\
9 & video of you that is very embarrassing & 88.41\% & 0.53 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingvideoofyoudoingsomethingembarrassingcombined}\\
10 & photo of you at home & 87.50\% & 0.60 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookphotosofyou(withaninward-facingcamera)athomecombined}\\
 & \vdots & & & \\
64 & eye patterns (for eye tracking) & 40.51\%& 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\
65 & exercise patterns  & 38.66\% & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenhowandhowmuchyouexercisecombined}\\
66 & when you are happy or having fun  & 34.75\% & 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouwerehappyorhavingfuncombined}\\
67 & television shows watched & 30.20\% & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhattelevisionshowsyouwatchcombined}\\
68 & when you are busy or interruptible  & 29.50\% & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouarebusyorinterruptiblecombined}\\
69 & music on device  & 28.06\% & 1.43 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/copiedanduploadedmusicfromyourdevicecombined}\\
70 & your heart rate & 27.50\% & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourheartratecombined} \\
71 & age & 24.29\% & 1.43 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyouragecombined}\\
72 & language spoken & 15.86\% & 1.49 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedthelanguageyouwerespeakingcombined}\\
73 & gender & 15.00\% & 1.45 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourgendercombined}\\ 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients. For the complete list of all data types across all recipients, see Appendix \ref{sec:concerns-appendix}.}
\label{top10-table}
\end{center}
\end{table*}

@@

\subsubsection{Data Recipient}
A statistically significant difference in VUR exists between data shared with an application versus data shared with human recipients. On average, 42\% of participants stated that they would be ``very upset'' if their data was shared with only an application's servers, whereas the VURs for friends (70\%), work contacts (75\%), and the public (72\%) were almost double (Table \ref{recipient-vur}). A chi-square test indicated that these differences were statistically significant (Table \ref{recipient-chi}). However, these effect sizes were small: the largest effect was between work contacts and an app's server ($\phi=0.11$); while the VUR for sharing with work contacts was significantly higher than sharing with friends, the effect size was negligible ($\phi=0.004$).

@@

However, we do not claim that there are no distinctions between the friends, public, and work contact recipients. People are more comfortable sharing certain data types with certain human recipients. For instance, participants were significantly uncomfortable sharing if they were lying, nervous, or stressed to work contacts compared to the rest of the data recipients. Table \ref{all-vur} Appendix \ref{sec:concerns-appendix} shows the  complete VUR and rankings of all data types by recipient.

%People have been shown to behave differently, especially in the social arena \cite{dwyer2007trust, gross2005information, fogel2009internet}.

@@

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{| c | r | r |}
\hline
 Misbehavior &  Cubetastic3000 & Smartphone \\
 \hline
 \hline
 All & 58.79\% & 46.67\%\\
 \hline
Vibration & 14.81\%  &  6.14\%\\
Bluetooth & 44.12\%  &  19.86\%\\
Unmuted Call & 87.10\%  &  58.44\%\\
Screenshot & 52.78\%  & 55.74\%\\
Premium Calls/Texts & 86.49\%  &  91.94\%\\ 
\hline
\end{tabular}
\caption{VURs for the five questions about device misbehaviors described in Section \ref{sec:smartphones}, contrasting smartphones with the Cubetastic3000.}
\label{deviceVUR}
\end{center}
\end{table}

%\subsubsection{Device}
%Each participant answered two questions drawn from a set of five smartphone misbehaviors. To compare these misbehaviors with misbehaviors on the Cubetastic3000, we included these same five questions amongst the pool of 293 Cubetastic3000 scenarios, modifying the device in question. In this manner, all 1,782 participants received two smartphone questions, while 159 participants received at least one of these five Cubetastic3000 misbehavior questions. Across all participants, the VUR was 46.7\% (of 1,782) for smartphones, whereas the VUR increased to 58.8\% (of 159) for same misbehaviors on the Cubetastic3000. The VURs for both devices for these questions are in Table \ref{deviceVUR}.
%
%To ensure independence of observations, we performed a Mann-Whitney U test to compare participants' average VURs for the Cuebtastic3000 scenarios (i.e., 159 participants) to the remaining participants' average VURs for the smartphone scenarios (i.e., 1,623 participants). This difference is statistically significant ($U=108,664.0$ with $p<0.0005$) but the effect size is very small ($r=0.08$). Because of this, we did not further reduce our statistical power by separately comparing the five misbehaviors. We conclude that users are likely more wary of misbehaviors on wearables than smartphones, but this difference is likely negligible. The entire effect may be due to participants' familiarity with smartphones, and therefore this effect may disappear as they increasingly encounter more wearables.

%{\color {red} talk about the variance of the devices in general--was there more spread for the answers wrt smartphones, cubetastic, or were they kind of all the same?}

>>>>

\begin{abstract}
\indent\indent Wearable devices, or ``wearables,'' bring great benefits but also potential risks that could expose users' activities without their awareness or consent. We surveyed 1,782 Internet users about risks associated with the capabilities of popular wearable devices on the market in order to identify risks that users find concerning. Our study  relatively ranks a range of potential wearables data capture scenarios, investigates the impact of the recipient of the data on the perceived risk of data capture, and explores the landscape users' concerns associated with wearable devices (including privacy and security, but also including factors such as increased safety risk, changes in social behaviors, and how wearable devices are unfashionable). To our knowledge, this is the largest user-based experiment concerning security and privacy concerns regarding wearables. We hope that this work will inform the community about user perceptions of wearable devices and aid in the design of future user notification, permission management, and access control schemes for wearables.
\end{abstract}

@@

Wearable devices enable benefits ranging from a fitness-data inspired lifestyle to an virtual-object filled augmented reality. However, wearable devices also bring new potential privacy and security risks that could expose users' activities without their awareness or consent. Although wearable devices are still in their infancy, we have already seen manifestations of these risks. Fitbit's default privacy settings inadvertently exposed information about some of their users' sexual activity~\cite{Fitbit}. Public discomfort toward facial recognition caused Google to prohibit Google Glass applications from using facial recognition~\cite{GlassDetection}, but still resulted in tech hate crimes against its users ~\cite{1_russell_2014, 16_gross_2014}.

@@

The goal of this work is to motivate research on the still-malleable future of wearable interaction models to preserve privacy and security, which we found are the top user perceived risk associated with wearable devices.  Our survey of 1,782 Internet users contributes the following: %\\[-0.8cm]

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item We give a relative ranking of 72 potential wearables capture scenarios, which were inspired from the capabilities from the most popular wearable devices on the market at the time of the study. 
\item We study 4 possible data recipients to find that the recipient of the data contributes to the magnitude of overall perceived risk, but do not find statistically significant correlated factors of risk between data and recipient. 
\item We sketch a landscape of users' self-reported concerns regarding wearable devices and analyze responses using logistic regression models. 
\end{itemize}

\section{Methodology}
We designed a survey for our IRB-approved study to capture the general public's perception of wearables risks. To determine which of the current data capture scenarios were most concerning to users, we asked them to rate their level of concern for a handful of scenarios from a list of possible scenarios. This was intended to elicit their perception of the severity and impact of the risk. The format of this section was based on Felt {\it et al.}'s study of user perceptions of security and privacy risks with mobile devices~\cite{Felt}. To get a qualitative, unbounded measurement of what people thought the most common risk associated with wearables are, we ask our participants an open-ended question.

To obtain a representative list of scenarios, we examined the sensors, capabilities, permissions, and applications of the most popular wearable devices on the market. At the time of this study (August 2014) the most popular wearable devices included the Fitbit fitness tracker, which continuously monitors heartbeat, steps taken, and sleep patterns; the Pebble smartwatch, which can take pictures, send texts, show notifications from online, and push notifications to services; and Google Glass, which can take pictures, record video, and perform a subset Internet-based tasks such as search, reading emails, etc. These devices' capabilities and requested permissions were the basis for the list of possible data capture scenario questions used in this study, which we feel will be representative of what users are likely to encounter today.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item 2 reading comprehension questions
\item 6 questions regarding possible wearables scenarios
\item 1 open-ended wearables risk question
\item 14 demographic questions %\\[-.8cm]
\end{itemize}

To assist in user fatigue, gave our participants a randomly selected subset of wearables scenarios. The average survey completion time was 11.5 minutes, which includes time for answering questions that are omitted from this paper due to lack of participants' familiarity with specific wearables devices and a misguided attempt to directly compare smartphones and wearables. See Appendix \ref{sec:noquestions} for details.

\subsubsection{Comprehension Questions}
The beginning of the survey introduced participants to the ``Cubetastic3000,'' which was the basis for all questions on wearables risks. Because participants might be biased to specific companies (e.g., visceral reactions to Google Glass based on popular media stories), we framed our scenarios on a fictitious wearable. We highlighted the capabilities of this device and described use cases:

\begin{quotation}
{\it Imagine that you are the proud owner of the Cubetastic3000, a new, high-tech computing device designed to be worn on your head. Imagine that you wear this device all the time, because it is very lightweight, durable, and convenient.

The Cubetastic3000 has the capability to capture video, photos, audio, and biometrics (biological data about you, such as heart rate). Just like other devices, you can install third-party applications from an app store, and these applications can use the information from the Cubetastic3000.

With a wide range of applications, your device can do all sorts of things, such as:\\

@@

To guarantee that participants had understood its capabilities, we asked two multiple-choice comprehension questions and filtered out responses from participants who did not answer both questions correctly.

\subsubsection{Wearable Scenarios}
We presented scenarios involving data captured by the Cubetastic3000 and asked participants to rate how upset they would be if a particular type of data (e.g., how much you exercise) was shared without permission with a particular recipient (e.g., work contacts). The purpose for using this question format was to determine how upset participants would be if data were inappropriately shared, and the extent to which their reactions were based on the data type and recipient. Responses were reported on a 5-point Likert scale (from ``indifferent'' to ``very upset'').. Figure \ref{fig:prompt}) shows an example.  Specifically, questions were of the form:

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{images/prompt.png}
	\caption{An example of a wearable scenario question participants saw while taking the survey.}
	\label{fig:prompt}
\end{figure}

We combined 72 data types (<data>) with 4 data recipients (<recipient>) to form an initial pool of 288 questions.We randomly pick 6 wearables scenarios from a list of 288 scenarios, and also display them to the user in a random order. Our 288 scenarios are the result of permuting 72 scenarios with 4 recipients, which can be seen in Table \ref{all-vur}. We clarified for our participants that ``app'' meant the the app's server, and that the data was not shared with anyone else. Each participant answered 6 questions that were randomly drawn from this pool.

\subsubsection{Additional Questions}
The exit portion of the survey contained demographic questions asking for age, gender, and education. We then asked about wearables ownership so we could control for prior exposure. An open-ended question on what would be the most likely risks associated with wearable devices was included to catch user concerns more broadly. To avoid biasing the open-ended question, we asked before concluding with the 10-question Internet Users' Information Privacy Concerns (IUIPC) index~\cite{malhotra2004internet}, which was included so we could control for participants' general privacy attitudes. However, we do realize that we asked this open-ended question after we had exposed our participants to a variety of wearables scenarios, which may have heightened their awareness of the possible risks. We talk about this more in Section 4.

@@

\section{Results}
In this section, we present participants' responses to the various data-sharing scenarios, and discuss how and which various factors contributed to their risk perceptions. We had at least 141 responses per data type, 2,779 per recipient, and 35 responses per each unique data type/recipient combination. We conclude the section with participants' self-reported wearables concerns.\\

\noindent {\bfseries Data Type}
Based on our statistical models (later reported in Section \ref{sec:regression}), we observe that the largest effect on participants' VURs stemmed from the data being shared, rather than  with whom the data is shared. The most and least concerning data types are listed in Table \ref{top10-table}.

@@

Although certain data is considered unanimously upsetting to have shared, it is interesting to note that no data was considered unanimously non-upsetting to have shared, nor were there any data types that evoked strong disagreement between participants (i.e., bimodal). Generally, the average concern magnitude was inversely correlated with the standard deviation, which suggests the presence of ceiling effects for the most concerning data types. For the complete ranked list of data types in this study, see Table \ref{all-vur}.\\

\begin{table*}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | r |c |}
\hline
Rank & Data & VUR & $\sigma$ & Distribution \\
\hline
1 & video of you unclothed & 95.97\% & 0.31 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouunclothedcombined} \\
2 & bank account information & 95.91\% & 0.35 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourbankaccountinformationcombined}  \\
3 & social security number & 94.84\% & 0.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyoursocialsecuritynumbercombined}\\
4 & video entering in a PIN at an ATM & 92.67\% & 0.47 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouenteringinyourPINatanATMcombined}\\
5 & photo of you unclothed & 92.59\% & 0.46 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookaphotoofyouunclothedcombined}\\
6 & photo of you that is very embarrassing & 91.39\% & 0.55 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingphotoofyoudoingsomethingembarrassingcombined}\\
7 & username and password for websites & 89.55\% & 0.62 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourusernameandpasswordforwebsitescombined}\\
8 & credit card information & 88.98\% & 0.56 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourcreditcardinformationcombined}\\
9 & video of you that is very embarrassing & 88.41\% & 0.53 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingvideoofyoudoingsomethingembarrassingcombined}\\
10 & photo of you at home & 87.50\% & 0.60 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookphotosofyou(withaninward-facingcamera)athomecombined}\\
 & \vdots & & & \\
64 & eye patterns (for eye tracking) & 40.51\%& 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\
65 & exercise patterns  & 38.66\% & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenhowandhowmuchyouexercisecombined}\\
66 & when you are happy or having fun  & 34.75\% & 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouwerehappyorhavingfuncombined}\\
67 & television shows watched & 30.20\% & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhattelevisionshowsyouwatchcombined}\\
68 & when you are busy or interruptible  & 29.50\% & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouarebusyorinterruptiblecombined}\\
69 & music on device  & 28.06\% & 1.43 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/copiedanduploadedmusicfromyourdevicecombined}\\
70 & your heart rate & 27.50\% & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourheartratecombined} \\
71 & age & 24.29\% & 1.43 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyouragecombined}\\
72 & language spoken & 15.86\% & 1.49 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedthelanguageyouwerespeakingcombined}\\
73 & gender & 15.00\% & 1.45 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourgendercombined}\\ 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10-table}
\end{center}
\end{table*}

@@

\noindent {\bfseries Data Recipient}
A statistically significant difference in VUR exists between data shared with an application versus human recipients. On average, 42\% of participants stated that they would be ``very upset'' if their data was shared with only an application's servers, whereas the VURs for friends (70\%), work contacts (75\%), and the public (72\%) were almost double (Table \ref{recipient-vur}). A chi-square test indicated that these differences were statistically significant (Table \ref{recipient-chi}). However, these effect sizes were small: the largest effect was between work contacts and an app's server ($\phi=0.11$); while the VUR for sharing with work contacts was significantly higher than sharing with friends, the effect size was negligible ($\phi=0.004$).

The statistical significance arises for two distinct reasons. Firstly, sharing data only with an application's server carries less social impact. And for our participants, it may seem that it is shared with less people. Additionally, there is a class of data which may be considered odd for a human to know, but completely normal for a wearable device to know (e.g., it's okay if your fitbit knows when you sleep, but maybe less so for your friends).

@@

However, we do not claim that there are no distinctions between the friends, public, and work contact recipients. People are more comfortable sharing certain data types with certain human recipients. For instance, participants were significantly uncomfortable sharing if they were lying, nervous, or stressed to work contacts compared to the rest of the data recipients.Table \ref{all-vur} shows the  complete VUR and rankings of all data types by recipient.\\

@@

\noindent {\bfseries Open-ended Concerns}
We captured participants' reactions to wearable devices as a whole by asking the following open-ended question:
